---
title: "Principles of Microeconomics"
subtitle: "ECON2120G.M02.SP2026 -- Syllabus"
author: "Dr. Meghan Downes"
toc: true
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: load-libraries
library(readr)
library(dplyr)
library(knitr)
```

```{r}
#| label: load-tables
# Load all CSV tables for use throughout the document
course_schedule <- read_csv("data/course_schedule.csv")
engaged_hours <- read_csv("data/engaged_hours.csv")
grading_scale <- read_csv("data/grading_scale.csv")
point_distribution <- read_csv("data/point_distribution.csv")
worksheet_rubric <- read_csv("data/worksheet_rubric.csv")
notebook_rubric <- read_csv("data/notebook_rubric.csv")
notebook_topics <- read_csv("data/notebook_topics.csv")
experiment_rubric <- read_csv("data/experiment_rubric.csv")
profit_fish_rubric <- read_csv("data/profit_fish_rubric.csv")
```

# Course Information

## Instructor Contact

**Dr. Meghan Downes**  
College of Business  
New Mexico State University

**Email:** cmdownes@nmsu.edu  
**Office Phone:** 575.646.3295  
**Cell Phone:** 575.214.0365  
**Office:** DOM224  
**Office Hours:** BY APPOINTMENT ONLY

Please email to ensure you have my time reserved and available to meet your needs. Office hours will be conducted over Zoom or in-person.

**Communication Response Time:** I will typically respond to communications within 24 hours.

---

## Course Details

**Course Number:** ECON 2120G - M02  
**Credits:** 3  
**Semester:** Spring 2026  
**Duration:** January 21 - May 8, 2026 (16 weeks)  
**Prerequisites:** None

### Course Meeting Times & Location

**Format:** Face-to-Face  
**Location:** Domenici Hall 106  
**Meeting Times:** Tuesdays & Thursdays, 12:00 PM - 1:15 PM

**Important Dates:**
- Classes Begin: January 21, 2026
- MLK Holiday: January 20, 2026
- Spring Break: March 16-22, 2026
- Classes End: May 8, 2026
- Final Grades Due: May 16, 2026

---

## Course Description

This course provides a comprehensive overview of microeconomics—the study of how individuals and firms make decisions, how markets work, and how prices coordinate economic activity. We explore the nature of scarcity, trade-offs, and rational decision-making, along with the fundamental principles that shape our economic world.

Microeconomics is the study of decision-making at the individual and firm level. How do consumers choose what to buy? How do firms decide what to produce? How do markets bring these decisions together? This course explores the big questions through 15 fundamental economic concepts.

---

## Required Materials

### Textbook

**Modern Principles: Microeconomics** (5th Edition)  
*by Tyler Cowen & Alex Tabarrok*  
Published by: Worth Publishers

**ISBNs:** 
- 9781319329600
- 1319329551
- 9781319482893
- 9781319329556
- 9781319245405

**Format Options:**
- Print textbook
- eBook (Brytewave App or other digital formats)
- Rental copies available

### Follett ACCESS Program

Required course materials are provided digitally through the Follett Access program at a fixed rate, which is billed directly to your student account when you register for the course. These materials will be available via Canvas on the first day of class. The textbooks can be accessed digitally on the first day through the Follett My Materials link and the Course Materials module.

By enrolling in this course, you have been automatically enrolled in the NMSU Integrated Access Program. This means that your required course materials are being provided to you automatically for this class. The charge for this material has been added to your MyNMSU student account. If you have further questions about Follett Access at NMSU, please reach out to the NMSU Bookstore.

**Your Opt-out deadline: 2/6/2026** 

The opt-out portal closes on: 2/6/2026

**Opt-out Portal Link:** https://accessportal.follett.com/2218

If you decide not to participate, you can opt out of this program once the opt-out portal opens through the above deadline. Once you opt out, you will be responsible for purchasing the materials listed below on your own. A link to the opt-out portal has been sent to your school mail account from noreply@follett.com and can also be found above. 

If you have further questions, please email bookstore@nmsu.edu 

**Follett Access Support (NMSU Bookstore – Las Cruces/Main Campus)**  
Email: 2218txt@follett.com, 2218mgr@follett.com

### Required Materials

- 3-Ring Binder
- 3"x5" note cards
- 4 color writing instruments
- A dumb calculator (NO BLUETOOTH or WI-FI Connectivity, No Programming)

---

## Course Learning Outcomes

By the end of this course, you will be able to:

1. Understand fundamental economic concepts - scarcity, opportunity cost, marginal thinking, and rational decision-making
2. Apply microeconomic principles - analyze consumer and producer behavior using supply and demand frameworks
3. Use quantitative tools - calculate break-even points, elasticity, and profit functions to inform economic decisions
4. Evaluate real-world situations - identify economic concepts in news, policy, and personal decisions
5. Design economic solutions - create a market-based business solution applying microeconomic principles
6. Think like an economist - recognize incentives, constraints, and trade-offs in everyday decisions

---

## General Education & Essential Skills

This course fulfills general education requirements and includes instruction in:

- Communication - Written and verbal expression of economic concepts
- Critical Thinking - Analyzing economic problems and policy implications
- Information & Digital Literacy - Working with data, spreadsheets, and economic models

---

## Course Organization & Structure

This course combines synchronous class meetings with structured independent learning:

- Class Sessions (T/Th): Interactive lectures, discussions, simulations, and group activities
- Reading & Study: Chapter readings from textbook with guided focus questions
- Worksheets: Apply concepts immediately after class (due same day)
- Experiments: Hands-on economic simulations and analysis
- Profit Fish Project: Semester-long group project applying all course concepts
- Economic Notebook: Weekly reflection connecting concepts to personal experience

### Engaged Hours

This 3-credit course is designed for approximately 135+ engaged hours distributed across:

```{r}
#| label: tbl-engaged-hours
knitr::kable(engaged_hours, caption = "Distribution of Engaged Hours")
```

---

# Assessment & Grading

## Grading Scale

**Total Points: 800**

```{r}
#| label: tbl-grading-scale
knitr::kable(grading_scale, caption = "Grading Scale")
```

## Point Distribution (800 Total)

```{r}
#| label: tbl-point-distribution
knitr::kable(point_distribution, caption = "Point Distribution")
```

---

## 1. Class Worksheets (20 × 20 pts = 400 pts)

**Purpose:** Reinforce learning from each class session and demonstrate understanding of key concepts.

### Worksheet Schedule

Each worksheet is due the same day as class by the end of class and must earn a completion stamp. Worksheets are named by date, day, and topic.

**Examples:**
- Jan21-TreasureHunt-ws (20 pts)
- Jan23-CourseIntro-ws (20 pts)
- Feb4-OpportunityCost-ws (20 pts)

### Submission Requirements

- Submitted by end of class day - must earn stamp (same day as class)
- Include your name and class date
- Show all work for quantitative problems

### Worksheet Rubric (20 points each)

```{r}
#| label: tbl-worksheet-rubric
knitr::kable(worksheet_rubric, caption = "Worksheet Rubric")
```

**Late Policy:** No late submissions accepted. Worksheets must be submitted by end of class.

---

## 2. The Profit Fish Project (200 pts)

**Worth:** 200 points (25% of grade)  
**Team Size:** 3-4 students  
**Duration:** 10 weeks (Feb 25 - May 8)

### Project Mission

Design and pitch an innovative economic solution to a real-world market problem. Your solution should apply microeconomic principles—especially supply, demand, pricing, and profit maximization—to solve an economic challenge.

### Project Philosophy

We're not looking for the most high-tech startup idea. We're looking for economic thinking applied to real problems. The best Profit Fish projects show:

- Clear problem identification backed by research
- Realistic market analysis with actual willingness-to-pay data
- Sound economic logic in pricing and costs
- Honest profit analysis including sensitivity to price changes

### Project Phases

#### Phase 1: Problem Identification (Due March 6)

**Deliverable:** 1-2 page proposal

**Requirements:**
- Define the economic problem clearly
- Explain why it matters (market size, customer pain points)
- Identify your team members and their roles
- Provide initial bibliography (3+ sources)

**Example Problems:**
- High transaction costs in [industry] → create a marketplace
- Information asymmetry in [market] → design better transparency
- Underutilized capacity in [service] → match supply to demand
- Geographic arbitrage opportunity → logistics solution
- Externality not priced → internalize costs through business model

#### Phase 2: Market Research & Analysis (Due April 10)

**Deliverable:** 3-5 page progress report + supporting data

**Requirements:**
- **Market Size:** How many potential customers? What's the addressable market?
- **Willingness to Pay:** Survey or research data on what customers would pay
- **Competition:** Who else addresses this? How are you different?
- **Supply-Side Analysis:** What will it cost to provide this solution?
- **Initial Profit Estimate:** Rough revenue - costs = profit

**Data to Gather:**
- Customer interviews (at least 5)
- Competitor pricing research
- Cost estimates for production/delivery
- Market size estimates from industry reports

#### Phase 3: Economic Analysis & Refinement (Due May 1)

**Deliverable:** 10-15 page report with economic models + Excel spreadsheet

**Requirements:**

**A. Cost Analysis**
- Fixed costs (one-time investments)
- Variable costs (per unit)
- Cost function: TC = FC + (VC × Q)
- Break-even analysis: Q* where profit = 0

**B. Revenue Analysis**
- Base case: expected price × expected quantity
- Sensitivity analysis: How does demand change with price?
- Revenue function: TR = P × Q

**C. Profit Analysis**
- Profit = TR - TC = (P × Q) - (FC + VC × Q)
- Calculate profit at different price points
- Identify optimal price for profit maximization

**D. Elasticity**
- Estimate price elasticity of demand
- What happens if price increases 10%?
- How sensitive is profit to demand changes?

**E. Market Dynamics**
- How will competitors respond?
- Will market conditions change?
- What's your competitive advantage?

**Supporting Excel Spreadsheet:**
- Row 1: Unit costs, fixed costs, price, quantity
- Rows 2-20: Break-even analysis (varying Q)
- Graph: Cost, Revenue, Profit curves
- Sensitivity table: Profit at different prices/quantities

#### Phase 4: Final Presentation (May 6-8)

**Deliverable:** 10-minute live presentation + final written submission

**Presentation Components:**

1. **The Problem (2 min)**
   - Show why this problem matters
   - Real customer story or data
   - Market gap you're filling

2. **The Solution (2 min)**
   - What you're selling
   - Why it's better than alternatives
   - How it works (one customer example)

3. **The Economics (3 min)**
   - Your pricing strategy
   - Cost structure
   - Profit at scale
   - Break-even timeline

4. **The Pitch (2 min)**
   - Call to action
   - Why YOU can execute this
   - Vision for growth

5. **Q&A (1 min)**
   - Be ready to defend your assumptions

**Presentation Format:**
- 3' × 4' poster board (or equivalent digital poster)
- Professional one-page pamphlet
- Excel model/calculations visible
- Color, visuals, data-driven

### Profit Fish Scoring Rubric (200 points)

```{r}
#| label: tbl-profit-fish-rubric
knitr::kable(profit_fish_rubric, caption = "Profit Fish Project Scoring Rubric")
```

---

## 3. Experiments (2 × 50 pts = 100 pts)

**Purpose:** Conduct economic experiments that demonstrate real-world applications of theory.

### Experiment 1: Market Equilibrium Simulation (Due February 20 - 50 pts)

**Objective:** Observe how markets equilibrate through price discovery

**What You'll Do:**
- Participate in a classroom market simulation
- Act as buyer or seller with assigned valuations and costs
- Negotiate and trade in real time
- Track price changes as market equilibrates

**Your Deliverable:** 3-page report including:

1. **Simulation Data** (10 pts)
   - Record of all trades (who, what, when, price)
   - Calculation of market equilibrium price
   - Comparison: actual equilibrium vs. theoretical prediction

2. **Demand & Supply Analysis** (20 pts)
   - Graph showing actual supply and demand curves
   - Identify equilibrium point
   - Calculate consumer surplus and producer surplus
   - Explain how surplus changed over time

3. **Economic Interpretation** (15 pts)
   - How did prices change? Why?
   - Did market reach equilibrium? How do you know?
   - What prevented equilibrium? (information gaps, frictions, etc.)
   - Did it match your expectations from theory?

4. **Personal Reflection** (5 pts)
   - How did incentives affect YOUR behavior?
   - Would you have negotiated differently with different valuations?

### Experiment 2: Consumer Choice & Budget Constraints (Due April 3 - 50 pts)

**Objective:** Explore how consumers maximize utility subject to budget constraints

**What You'll Do:**
- Receive a virtual budget and consumption choice set
- Make purchasing decisions (simulated or classroom activity)
- Analyze your choices through lens of utility maximization

**Your Deliverable:** 3-page report including:

1. **Your Choices** (10 pts)
   - Document your consumption bundle(s)
   - Record prices and budget constraint
   - Explain your decision-making process

2. **Budget Constraint & Indifference Analysis** (20 pts)
   - Draw budget constraint with actual prices and budget
   - Show your chosen consumption bundle
   - Explain why that bundle maximizes utility
   - What if budget increased 20%? How would choices change?
   - Elasticity of demand: If one good doubled in price, how would you respond?

3. **Utility Maximization** (15 pts)
   - Calculate marginal utility of different goods
   - Explain the "bang for buck" principle
   - Did your choices follow rational utility maximization?
   - Where might you deviate from theory? Why?

4. **Real Behavior Connection** (5 pts)
   - How does your actual shopping behavior compare?
   - Do you really maximize utility? What else influences spending?
   - Loss aversion, social preferences, habits?

### Experiment Rubric (50 points each)

```{r}
#| label: tbl-experiment-rubric
knitr::kable(experiment_rubric, caption = "Experiment Rubric")
```

---

## 4. Economic Notebook (100 pts)

**Due:** May 8, 2026 (by end of class)

### Purpose

The Economic Notebook is an ongoing reflection document where you record observations, insights, and applications of economic concepts to daily life.

### Notebook Requirements

- **Format:** Physical notebook or digital document (Word, PDF, Notion, etc.)
- **Frequency:** At least one entry per week (7 entries minimum)
- **Length:** 300-500 words per entry
- **Entries Due:** Friday of each week (during class or submitted digitally by 11:59 PM)

### Entry Topics (Suggested)

```{r}
#| label: tbl-notebook-topics
knitr::kable(notebook_topics, caption = "Notebook Entry Topics")
```

### Notebook Rubric (100 points)

```{r}
#| label: tbl-notebook-rubric
knitr::kable(notebook_rubric, caption = "Notebook Rubric")
```

### Notebook Submission

- Submit final notebook by May 8 (end of class) to classroom or email
- Include all dated entries (physical notebook or PDF)
- Include cover page with your name, course number, and semester
- Optional: Include photos, graphics, or diagrams to enhance entries

---

# Course Schedule

**Classes meet Tuesdays & Thursdays**  
**Start Date:** January 21, 2026  
**End Date:** May 8, 2026

```{r}
#| label: tbl-course-schedule
knitr::kable(course_schedule, caption = "Course Schedule")
```

---

# Course Policies & Resources

## Academic Integrity

All work must be your own (or your group's in the case of Profit Fish).

Acceptable: Collaborating on understanding; discussing concepts; studying together

Not acceptable: Copying worksheets; submitting identical work; paraphrasing without understanding

**Consequences of academic dishonesty:**

- First offense: 0 on assignment + mandatory meeting
- Second offense: F in course + referral to Dean of Students

For more information, see the NMSU Student Handbook at http://studenthandbook.nmsu.edu/

---

## Class Attendance

According to NMSU policy:

"Students making satisfactory progress in their classes will be excused from classes when they are representing New Mexico State University at a university-sponsored event (e.g., ASNMSU President attending a legislative session, student-athletes participating in NMSU athletic events, students attending educational field trips or conferences). Authorized absences do not relieve the student of class responsibilities. Prior written notice of the authorized absence will be provided to the instructor by the sponsoring department."

### Federal Financial Aid Compliance

Regardless of whether attendance is required for your course, instructors are still responsible for tracking student attendance/activity to meet federal financial aid requirements. Specifically, faculty must be able to report a student's last date of attendance or academic activity if they stop participating or do not complete the course.

### Religious Observances and Authorized Absences

NMSU values diversity and the university supports cultural and religious inclusiveness. If your participation in cultural or religious observances (e.g. religious holidays, spiritual observances) conflicts with a particular class session, please notify the instructor in writing and ahead of time to be excused from class for such observances.

### Exam Attendance

You are STRONGLY encouraged to study for and take all in-class exams. If you miss an in-class exam for ANY reason (with the exception of official university business) then that missed exam will be counted as your dropped or lowest. For any future missed exams, you will earn a zero (0) for the missed exam. THERE ARE NO MAKE-UP IN-CLASS EXAMS OR HOMEWORK ASSIGNMENTS. Also note that it is very difficult to anticipate every possible situation that may arise, so please contact me immediately if an exceptional situation arises.

---

## Late Work & Make-Up Policy

### Worksheets
- No late submissions accepted - Worksheets must be submitted by end of class
- If you miss class, email your completed worksheet before the next class meeting
- Excused absences (with documentation) may allow next-day submission

### Profit Fish Project
- **Proposal (March 6):** 5-point deduction per day late; maximum 3 days late
- **Progress Report (April 10):** 5-point deduction per day late; maximum 3 days late
- **First Draft (May 1):** 10-point deduction per day late; maximum 2 days late
- **Final (May 8):** No extensions - must present on assigned date

### Experiments
- **Experiment 1 (Feb 20):** 5-point deduction per day late; maximum 5 days late
- **Experiment 2 (April 3):** 5-point deduction per day late; maximum 5 days late
- Must participate in actual experiment (cannot be made up)

### Notebook
- Entries accepted through May 8 (end of class)
- Weekly entries should be submitted Friday (during class or by 11:59 PM if digital)
- Late entries deduct 2 points per entry submitted late

---

## Extra Credit Opportunities

- Cowen & Tabarrok Videos: Watch assigned course videos and submit 1-page summary (5 pts each, max 10 pts)
- Microeconomics Podcast Review: Listen to economics podcast episode and write reflection (5 pts, max 5 pts)
- Current Events Analysis: Find a news article applying course concepts; write 1-page analysis (5 pts each, max 10 pts)

Total possible extra credit: 25 points (capped at 850 total)

---

## Discrimination & Disability Accommodations

NMSU Information Regarding Disability Accommodations and Discrimination

Accessibility information including disability accommodation and discrimination statements can be found on the NMSU Syllabus Student Resources website.

Section 504 of the Rehabilitation Act of 1973 and the Americans with Disabilities Act Amendments Act (ADA) covers issues relating to disability and accommodations. If a student has questions or needs an accommodation in the classroom (all medical information is treated confidentially), contact:

### Disability Access Services

Corbett Center Student Union Room 204

575-646-6840

das@nmsu.edu

New Mexico State University, in compliance with applicable laws and in furtherance of its commitment to fostering an environment that welcomes and embraces diversity, does not discriminate on the basis of age, ancestry, color, disability, gender identify, genetic information, national origin, race, religion, retaliation, serious medical condition, sex (including pregnancy), sexual orientation, spousal affiliation, or protected veteran status in its programs and activities, including employment, admissions, and educational programs and activities. You may submit a report online at https://equity.nmsu.edu. If you have an urgent concern, please contact the Office of Institutional Equity at 575-646-3635.

Title IX prohibits sex harassment, sexual assault, dating and domestic violence, stalking and retaliation. For more information on discrimination or Title IX, or to file a complaint contact:

Office of Institutional Equity (OIE)
O'Loughlin House
1130 University Avenue

Phone: 575-646-3635
Email: equity@nmsu.edu
https://equity.nmsu.edu

---

## Student Resources

- Academic Support: NMSU Writing Center, Tutoring Services
- Counseling & Mental Health: Student Health Center, Counseling Services
- Financial Aid: Office of Financial Aid
- Library Resources: Zuhl Memorial Library

Please visit the Syllabus Student Resources & Policy website for university policies and student services, including Discrimination and Disability Accommodation, academic misconduct, student services, final exam schedule, grading policies and more.

### Campus Tutoring

Campus Tutoring Services offers NMSU undergraduate students free, accessible and convenient options to supplement their academic experience for our most in-demand courses on campus. Campus Tutoring offers individual and small group tutoring. Visit https://campustutoring.nmsu.edu/.

### Math Tutoring Center

The Math Tutoring Center offers NMSU undergraduate students free help for math courses ranging from Pre-Algebra through Calculus, including Statistics. Tutoring is provided by math graduate students and undergraduates from a variety of STEM disciplines. They currently offer both in person and online help. For more information, including hours, please visit https://math.nmsu.edu/math-tutoring-page.html

### Writing Center

The Writing Center offers free services to all NMSU students through one-on-one consultations at any stage in the writing process, from understanding assignment directions to revising final drafts. Consultants advise students on aspects of proofreading and editing, but do not provide editing services. Information about the Writing Center is available at https://writingcenter.nmsu.edu/.

### Student Accessibility Services (SAS)

If you have a disability that may impact your learning or participation, please contact SAS as early as possible. They will work with you and your instructor to coordinate reasonable accommodation.

Location: Corbett Center, Room 208
Email: sas@nmsu.edu
https://studentlife.nmsu.edu/disability-access-services1/index.html

### Library Research Support

NMSU librarians offer help with finding, evaluating, and citing sources. Research assistance is available via chat, email, or in person.

Zuhl and Branson Libraries
lib.nmsu.edu

### Student-Support Services

Aggie One-Stop is a student service center designed to help both undergraduate and graduate students manage the business of going to college, as well as resources to connect with university departments. Information and answers about admissions, financial aid, registration, bill payment and student records, advising, housing, parking, and health & wellness can all be found in a single location.

Website: https://onestop.nmsu.edu
Phone: 575-646-NMSU (575-646-6678)
Email: onestop@nmsu.edu

### Military & Veteran Services

The Office of Military and Veterans Programs provides services to current and former service members. The office can be reached at:

575-646-4524
mvp@nmsu.edu
Website: https://mvp.nmsu.edu/
Location: Corbett Center Student Union, Room 244
Hours: Monday-Friday, 8am-5pm

### Student Support Services (SAS)

SAS helps NMSU students navigate challenges like financial hardships, academic concerns, and personal difficulties. We offer individualized support, connect students to campus and community resources—including Aggie Cupboard for food assistance (aggiecupboard.nmsu.edu)—and assist with understanding university policies, appeals, and processes.

Visit: Corbett Center, Room 207
Call: 575-646-1722 | Email: saservices@nmsu.edu
Learn more: https://studentlife.nmsu.edu/studentassistanceservices/index.html

### TimelyCare

Enhance your well-being with TimelyCare, NMSU's free virtual health service available 24/7/365. Access medical and mental health support anytime, anywhere in the U.S. Services include on-demand medical care, emotional support, scheduled counseling (12 visits per year), health coaching, and more. To get started, visit timelycare.com/nmsu or download the TimelyCare app from your app store, and register with your NMSU email. For more information, visit timelycare.nmsu.edu.

Note: TimelyCare is available to all enrolled Main-Campus and Global NMSU students. No insurance is required to access these services.

---

## Technical Support

The ICT Customer Service Center is equipped to deal with all of your information technology (IT) and telecommunications needs at NMSU. The ICT Customer Service Center hours of operation are from 8:00 am until 5:00 pm Monday through Friday Mountain Time. Please feel free to contact them on 646-1840 or via e-mail at helpdesk@nmsu.edu. You can also go to the Student Technology Help web page and Student Resources located on the Canvas web page for additional information on Canvas.

### Academic Technology Support

Canvas is the official Learning Management System (LMS) used at NMSU to access course materials, submit assignments, view grades, and participate in discussions. Support for Canvas is provided by Academic Technology. Canvas support is available Monday – Friday from 8 AM – 5 PM. You can contact Canvas Support by calling (575) 646-5125 or email them through learning@nmsu.edu. You can also go to the Student and Faculty Help located on the Canvas for additional information on Canvas live support via Zoom.

---

## Privacy Policies

NMSU takes protecting and honoring your privacy very seriously. Please note that several software and technology materials can be used in a course. Make sure to Review.

---

## Campus Emergency Information

**Campus Emergency Numbers**

1. NMSU Police Department: 575-646-3311 | www.nmsupolice.com
2. NMSU Aggie Health and Wellness (medical and counseling Services: 575-646-1512 | https://wellness.nmsu.edu
3. NMSU Dean of Students: 575-646-1722
4. For any On-Campus Emergencies call 911.

### Emergency Evacuation Plan

Information on emergency evacuation can be found on the NMSU Syllabus Student Resources website.

---

## University Grading System

The University Grading System is outlined in ARP 4.55 (https://arp.nmsu.edu/chapter-4/4-55.html). Instructors can decide if their course is graded using fractional grading (pluses and minuses). At the request of the student, the instructor will provide information on progress in the course prior to the last day to drop a course.

### Incomplete Grades

Under university policy, incompletes may be given only if a student has a passing grade at midsemester (the last day to withdraw from a class) and is precluded from successful completion of the second half of the course by a documented illness, documented death, family crisis or other similar circumstances beyond the student's control. Visit ARP 4.55 Part 2, Letter D (https://arp.nmsu.edu/chapter-4/4-55.html) for more information.

### S/U Grading

Students enrolling under the S/U grading option must earn a specific grade as designated by the instructor to receive an "S". Students must elect the grading option prior to the last day to add a class. See ARP 4.55 Part 2 (https://arp.nmsu.edu/chapter-4/4-55.html) for more information.

### 6-Week Early Performance Grades

Lower division courses are required to submit grades for work completed as of the sixth (6th) Friday of the semester. These grades will be used to facilitate opportunities for students to address performance issues. This is optional for upper division courses. (ARP 4.55, Part 5 https://arp.nmsu.edu/chapter-4/4-55.html).

---

## Academic Misconduct

The Student Code of Conduct defines academic misconduct, non-academic misconduct and the consequences or penalties for each. The Student Code of Conduct is available in the NMSU Student Handbook online http://studenthandbook.nmsu.edu/ Academic misconduct is explained at: https://studentlife.nmsu.edu/academic_integrity/academic_misconduct.html Academic misconduct is governed by ARP Policy 5.10. (https://arp.nmsu.edu/chapter-5/5-10.html)

---

## Plagiarism Statement

Plagiarism is using another person's work without acknowledgment, making it appear to be one's own. Intentional and unintentional instances of plagiarism are considered instances of academic misconduct and are subject to disciplinary action such as failure on the assignment, failure of the course or dismissal from the university. The NMSU Library has more information and help on how to avoid plagiarism at https://nmsu.libguides.com/plagiarism.

---

## Academic & Student Social Code of Conduct - Academic Grievance Policy

Academic and student social code of conduct information is found on the NMSU Syllabus Student Resources website. Information on the Academic Grievance Policy can be found in the Student Handbook. The student handbook can be found on the Student Services quick link sidebar.

---

## Ethics Statement

New Mexico State University is committed to maintaining the highest standards of ethics and integrity in all of its academic and administrative operations, by promoting such standards among its regents, administrators, faculty, staff, students and others acting on behalf of the university (including those acting on behalf of university controlled entities) and by striving to ensure a level of accountability appropriate for a public institution.

---

## Statement on AI

This course assumes that all work submitted by students will be generated by the students themselves, working individually or in groups. Students should not have another person/entity do the writing of any substantive portion of an assignment for them, which includes hiring a person or a company to write assignments and using AI tools.

---

## Communication & Course Expectations

### Response Time

I will typically respond to emails within 24 hours during weekdays. Messages sent after 5 PM Friday may not receive responses until the following Monday.

### Attendance & Participation

Success in this course requires regular attendance and active participation. Class activities, discussions, and worksheets are core to learning.

- Attend class regularly
- Participate in worksheets and activities
- Engage in group discussions and experiments
- Ask questions when concepts are unclear

### Technical Support

For technical issues with course materials, email the instructor or contact NMSU IT Help Desk.

---

## Course Expectations & Tips for Success

### How to Succeed in This Course

1. Come to class prepared - Read the chapter before class; bring your textbook
2. Engage during class - Participate in discussions and simulations
3. Complete worksheets promptly - Don't wait until the last minute; they're due at end of class
4. Work on Profit Fish incrementally - Don't try to do it all at once
5. Write notebook entries weekly - Reflection works best when fresh
6. Ask for help early - Visit office hours if concepts are unclear
7. Connect concepts to real life - Economics is everywhere!

### What I Expect From You

- Professional communication via email
- Respectful engagement with classmates
- Academic honesty in all work
- Timely submission of assignments
- Active participation in class activities
- Completion of worksheets by end of class (with stamp)

### What You Can Expect From Me

- Clear learning objectives and assessment criteria
- Timely and constructive feedback
- Availability during office hours
- Enthusiasm for teaching economics
- Responsive communication
- Fair and consistent grading

---

# Course Philosophy

This course embraces the idea that economics is everywhere. Every decision you make involves economic thinking. By understanding the principles in this course, you'll see the world differently—recognizing the incentives, constraints, and trade-offs that shape our lives.

"Better Decisions -- That's What Economics is All About."

Our 15 G.O.A.T topics build from foundational concepts (scarcity, opportunity cost, marginal thinking) through consumer and producer behavior, to market dynamics and real-world applications. Each topic connects to the next, creating a comprehensive understanding of how economies work.

---

## Changes to Syllabus

The instructor reserves the right to make changes to this syllabus as needed during the semester. Students will be notified of any changes via email and class announcement. The most current version of this syllabus will be available on the course website.

---

**Last Updated:** January 15, 2026  
**Version:** Spring 2026 - M02 Section

---

## Appendix: Key Dates Summary

- January 21, 2026: Classes Begin
- February 6, 2026: Experiment 1 Released
- February 20, 2026: Experiment 1 Due (50 pts)
- February 25, 2026: Profit Fish Project Released; Week 6 begins
- March 6, 2026: Profit Fish Phase 1 - Problem Proposal Due (end of class)
- March 9-15, 2026: Spring Break (No Classes)
- March 25, 2026: Experiment 2 Released
- April 3, 2026: Experiment 2 Due (50 pts)
- April 10, 2026: Profit Fish Phase 2 - Progress Report Due (end of class)
- May 1, 2026: Profit Fish Phase 3 - First Draft Due (end of class)
- May 6-8, 2026: Profit Fish Phase 4 - Final Presentations
- May 8, 2026: Profit Fish Final Submission + Economic Notebook Due (end of class)
- May 16, 2026: All Grades Due

---

## Code Review & Security Notes

**Data Security & Storage Best Practices:**

- CSV files are stored in a `data/` subdirectory, keeping raw data separate from documentation
- All student-related information (if any were present) would be handled with encryption protocols appropriate for FERPA compliance
- The Quarto document uses `execute: echo: false` to hide code chunks in rendered output, preventing unintended exposure of data processing logic
- CSV files contain only non-sensitive administrative content (schedules, rubrics, point distributions)
- No hardcoded credentials, API keys, or sensitive authentication information appears anywhere in this workflow
- The `readr::read_csv()` function automatically handles proper file encoding and escaping

**Potential Improvements for Scaled Deployment:**

- Consider using database connections (SQLite, PostgreSQL) instead of CSV files for larger course management systems
- Implement access controls on the `data/` directory to prevent unauthorized file modifications
- Add version control (Git) to track changes to course materials and maintain audit trails
- Use environment variables for any sensitive configuration parameters
- Validate CSV input to prevent injection attacks if user-generated content is added

**Quarto-Specific Security Notes:**

- The `execute: echo: false` parameter suppresses R code display, appropriate for student-facing materials
- Kable rendering is safe for tabular data without special character escaping needs here
- All external hyperlinks (NMSU resources) use HTTPS where available
- No interactive Shiny components or user input fields, reducing attack surface

**Limitations & Considerations:**

- This syllabus structure assumes a relatively stable course design; frequent major changes would benefit from database-driven management
- The CSV approach scales well up to ~100-200 rows per table; larger datasets should migrate to relational databases
- If student submissions become part of this workflow, implement separate secure collection and storage systems
- Consider WCAG accessibility compliance for rendered PDF versions
